Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-06-18 vertical-thermal-review.Rmd
@teixeirak updated this sentence
Not sure if R2 read it as just respiration instead of Q10 but I added this sentence, let me know if it makes sense to the comment:

R2's comment: we currently lack mechanistic understanding of how and why the temperature sensitivity of respiration varies across the forest vertical profile'. Other than that it varies according to the prevailing temperature difference? I think the temperature difference between upper and lower canopy positions, the theme of this review, should be mentioned in regard to this

Similar to photosynthesis, respiration tends to be higher in upper-canopy sun leaves, but its temperature sensitivity (*Q~10~* or *E~0~*) shows no definite trend along the vertical gradient.
Specifically, the temperature sensitivity of respiration can be constant within vertical profiles and in seedling sun vs. shade leaves<!--[<!--@carterExperimentalWarmingTropical2021-> @weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007 <!--@zaragoza-castellsClimatedependentVariationsLeaf2008; @xuSeasonalVariationTemperature2006->]-->, greater in upper-canopy leaves<!--[@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996]-->, or greater in the lower canopy <!--[@griffinCanopyPositionAffects2002]--> (Table `r table_leaf_metabolism`). 
This variation may be attributable to forest type, leaf traits and age [e.g., greater *Q~10~* in younger leaves, @zhouLeafageEffectsTemperature2015], or acclimation to high temperature that decreases *Q~10~* [@slotGeneralPatternsAcclimation2015; @carterExperimentalWarmingTropical2021]. 
Overall, however, we currently lack mechanistic understanding of how and why the temperature sensitivity of respiration varies across the forest vertical profile.

>>>>

Similar to photosynthesis, respiration tends to be higher in upper-canopy sun leaves, but its temperature sensitivity (*Q~10~* or *E~0~*) shows no definite trend along the vertical gradient.
Specifically, the temperature sensitivity of respiration can be constant within vertical profiles and in seedling sun vs. shade leaves<!--[<!--@carterExperimentalWarmingTropical2021-> @weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007 <!--@zaragoza-castellsClimatedependentVariationsLeaf2008; @xuSeasonalVariationTemperature2006->]-->, greater in upper-canopy leaves<!--[@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996]-->, or greater in the lower canopy <!--[@griffinCanopyPositionAffects2002]--> (Table `r table_leaf_metabolism`). 
This variation may be attributable to forest type, leaf traits and age [e.g., greater *Q~10~* in younger leaves, @zhouLeafageEffectsTemperature2015], or acclimation to high temperature that decreases *Q~10~* [@slotGeneralPatternsAcclimation2015; @carterExperimentalWarmingTropical2021]. 
Overall, although *Q~10~* is influenced by the temperature difference between upper and lower canopy positions, we currently lack mechanistic understanding of how and why the temperature sensitivity of respiration varies across the forest vertical profile.
